master thesis

### 2018-03-30 Paper - Latex/Paper - Skizze - v4.tex
Thesis - Result table ....

  Here we test in how far it is possible to use the Open Tree of Life (OTL), a synthesis of 
    phylogenetic trees on a backbone taxonomy (resulting in unresolved nodes), to expand available 
    information via phylogenetic trait prediction. We use the Global Biotic Interactions (GloBI) 
    database to categorise 25,992 and 34,879 species as parasites and free-living, respectively and 
    predict states for over $\sim$2.3 million (97.34\%) leaf nodes without state information.

@@

  % This study focuses on the ancestral state reconstruction of parasitism in the tree of life of
  %   Eukaryota. We predict unknown states of species and estimate origins and losses of parasitism. \\
  % The challenge here is the size of the tree and the little information about it. \\
  % Such a large phylogenetic tree does not completely exist and therefore we work with a synthesis 
  %   tree of OTL \cite{Hinchliff2015} which is highly multifurcated. \\
  % For the 2,535,437 leaf nodes we could not gather much data. From the GloBI database 
  %   \cite{Poelen2014} which we used, we could only collect 25,992 parasitic and 34,879 free-living 
  %   species. It follows that we have $\approx 2.4\%$ state information. \\
  % So far, especially small scale or highly manual studies have been carried out. In this scale, it 
  %   requires different data sources to be interconnected. \\
  % We performed an analysis of existing algorithms and selected a Sankoff maximum parsimony algorithm 
  %   using the R package \textit{Castor} \cite{Louca2017}. \\
  % Nevertheless, the results are convincing and even though this is a purely computational approach 
  %   which did not include human experts input, results coincide with prior knowledge. Also regarding 
  %   the number of events, our estimates coincide with previous results by human experts, e.g. the 
  %   study by Weinstein and Kuris \cite{Weinstein2016}. \\

@@

  To accomplish this task, next to the phylogenetic tree information about the current species' 
    states is needed. \\
  Most of the largest interaction databases are offline or outdated. For example: IWDB (Interaction 
    Web Database) \cite{IWDB2003}, Webs on the Web \cite{WOW2004}, Animal Diversity Web 
    \cite{Myers2003} and ecoweb \cite{Cohen2010}. \\
  We use the interaction database Global Biotic Interactions (GloBI) \cite{Poelen2014} because it is 
    including most of the known databases and is still growing actively \cite{Poelen2014}. \\
  The data in GloBI are stored as interactions e.g.  species A parasitize species B. From this we 
    conclude that species A is parasitic and species B free-living. \\
  From this data, we could specify $\sim$2.3 million leaf nodes 34,879 as free-living and 25,992 
    as parasitic.

@@

  The objectives of this work are the following points: (1) Find a suitable ancestral state 
    reconstruction method. (2) Accomplish reconstructing on the Eucaryotic synthesis tree of OTL. \\
  The goal of Point 1 is to evaluate the possible methods based on a simulation of our data 
    situation. \\
  The Sankoff algorithm implemented by Louca et al. is the best in our comparisons. Therefore, 
    point 2 consists of reconstructing the ancestral states and predicting the unknown leaf states 
    with the aid of this algorithm. And then perform an evaluation of the results.

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%--------------------------------------------------------------------------------------------------- 
\chapter{Aims}
  The aim of this thesis is the application of maximum parsimony algorithms to non-binary trees and 
    very large data sets to predict unknown leaf node states. \\
  We discuss different ancestral state reconstruction methods and test the most appropriate ones on 
    our data. Hereby, we want to find out how far it is possible to use the Open Tree of Life (OTL) 
    to expandavailable information via phylogenetic trait prediction. As a basis for known states, 
    we use the Global Biotic Interactions (GloBI) database, which cathegorises 34,879 and 25,992 
    species as free-living and parasites. This data is used to perform an ancestral state 
    reconstruction and predict states for over $\sim$2.3 million (97.34\%) leaf nodes without state 
    information.

@@

    Analogous to the missing leaf node states, generalized linear models with binomial regression are 
      compared according to their residuals, based on the "true" / "false" prediction. \\
    Again, in order to compare models of different complexity, the BIC (Bayesian Information Criterion) 
      values are calculated in addition to the residuals. \\
    For all these calculations, the following R functions are used: \textit{glm()}, \textit{anova()} 
      and \textit{BIC()}.

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Poisson regression of the multifurcation}
      For a regression analysis of the multifurcation first set up a simple generalized linear model and 
        then output the intercept of it:
      \lstset{language=R}
      \begin{lstlisting}[gobble=6]
        kingdom.furc.mod <- glm(multifurc~kingdom, 
                                    data=inner.taxa,
                                        family="poisson")
        # get the intercept:
        summary(kingdom.furc.mod)
      \end{lstlisting}
      Since the multifurcation can also be 0, the intercept describes the expected mean value and thus, 
        if it is greater than 0, it says that we have a multifurcation: $2.62 > 0$ $\Rightarrow$ there 
        is a multifurcation. \\
      We next compared the intercept of the different kingdoms and found that
        $$1.22 \text{ (Fungi intercept)} > -0.11 \text{ (Metazoa intercept)} > -0.09 \text{ (Chloroplastida intercept)}$$
        % Coefficients:
        %                                         Estimate Std. Error z value Pr(>|z|)
        % (Intercept)                             2.624897   0.003641  721.00   <2e-16
        % kingdom['ott352914', 'Fungi']           1.224650   0.004054  302.10   <2e-16
        % kingdom['ott361838', 'Chloroplastida'] -0.090993   0.003960  -22.98   <2e-16
        % kingdom['ott691846', 'Metazoa']        -0.114873   0.003732  -30.78   <2e-16
        that means the strength of the multifurcation \colorbox{red}{is the same/has the same size 
        sorting}. \todo{kann das sein?} \\

      We investigated several models that could describe multifurcation: \\
      In doing so, we allowed the different influence of the taxa and the heights and depths of a node 
        to be included. From this we got 9 times 4 models of different complexity levels (see first row 
        of table \ref{table:BIC multifurcation}). \\
      We first compared the models based on their residuals. These give us the error of the model. If 
        the value is small, our data will be well modeled. These values are meaningful only for the same 
        complexity, therefore we then calculated the BIC (Bayesian information criterion) values of the 
        models and used them to evaluate the models. \\
      % \lstset{language=R}
      % \begin{lstlisting}[gobble=6]
      %   anova(kingdom.furc.mod, phylum.furc.mod, class.furc.mod, 
      %                 order.furc.mod, family.furc.mod, test="LRT")
      %   BIC(kingdom.furc.mod)
      %   ...
      % \end{lstlisting}
      The deviance table with the residuals can be found in the appendix (table 
        \ref{table:Residuals multifurcation}). Table \ref{table:BIC multifurcation}. lists the BICs.

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & 8273333 & \cellcolor{green!15}7937828 & \cellcolor{green!20}7842157 & \cellcolor{green!30}7644249 \\
            multifurc $\sim$ taxa & 8257680 & \cellcolor{green!15}7922207 & \cellcolor{green!20}7826490 & \cellcolor{green!35}7574154 \\
            \hline
            multifurc $\sim$ taxa + depth & 8273318 & \cellcolor{green!15}7934322 & \cellcolor{green!20}7839364 & \cellcolor{green!35}7539999 \\
            multifurc $\sim$ taxa + max.height & \cellcolor{green!15}7993515 & \cellcolor{green!25}7749121 & \cellcolor{green!30}7661817 & \cellcolor{green!40}7416211 \\
            multifurc $\sim$ taxa + min.height & 8251211 & \cellcolor{green!20}7875521  & \cellcolor{green!25}7778327 & \cellcolor{green!35}7516883 \\
            multifurc $\sim$ taxa + mean.height & \cellcolor{green!20}7825417 & \cellcolor{green!30}7644249 & \cellcolor{green!35}7572474 & \cellcolor{green!45}7340741 \\
            \hline
            multifurc $\sim$ taxa * depth & 8235932 & \cellcolor{green!20}7836755 & \cellcolor{green!25}7757688 & \cellcolor{green!45}7383808 \\
            multifurc $\sim$ taxa * max.height & \cellcolor{green!15}7963438 & \cellcolor{green!30}7693555 & \cellcolor{green!30}7614820 & \cellcolor{green!45}7335338 \\
            multifurc $\sim$ taxa * min.height & 8214030 & \cellcolor{green!20}7808940 & \cellcolor{green!30}7690618 & \cellcolor{green!45}7336627\\
            multifurc $\sim$ taxa * mean.height & \cellcolor{green!25}7768360 & \cellcolor{green!35}7536296 & \cellcolor{green!50}7484953 & \cellcolor{green!50}7206369 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values of the multifurcation models. \\
          These models are created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values.}
        \label{table:BIC multifurcation} 
      \end{table}
      % * Residuals: Fehler - wieviele Werte sind nicht gut modelliert. (umso kleiner umso besser - grün) \\
      Within the every complexity class it can be seen that the mean height gives the best additional factor. \\
      Despite higher complexity, the BIC values are getting smaller from model to model, meaning that 
        the finest model available here is also the best one of these. Lower taxa than orders (e.g. 
        family) are computationally too expensive to calculate. \\
      The model \textit{multifurc $\sim$ order * mean.height} turns out to be the best of our models, 
        whereby it is possible that e.g. \textit{multifurc $\sim$ family * mean.height} is better.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Binomial regression of the unknown state information}

      Next to the problem of the multifurcation of the tree is the less of interaction data we have for 
        the species. For the ancestral state reconstruction, we need information about the states 
        (free-living or parasite) in the leaf nodes. \\
      The eukaryotic synthesis tree has 293,463 leaf nodes. The GloBI database has 5,346,414 interactions 
        (at 29.01.2018). Out of this data we got 51,337 distinct free-living species and 47,332 
        distinct parasite species for the whole tree of life. From the Eukaryota we could determine 
        25,992 and 34,879 species as parasites and free-living. With 2,293,463 leaf nodes we still have 
        about 97.34\% unknown leaf nodes. In the discussion, section 
        \ref{sec:discussion - data situation}, we will talk about ways to extend this data. \\

      We also compared different models in terms of their BICs (Table: \ref{table:BIC unknown information}). 
        The Residuals are not very meaningful here, since all models have different complexities. % /dimensions
        For the sake of completeness, the associated deviance tables are located with the residuals in 
        the appendix \ref{sec:Residuals unknown information}.

      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & \cellcolor{green!15}545799 & \cellcolor{green!35}500004 & \cellcolor{green!45}485121 & \cellcolor{green!45}484681 \\
            \hline
            multifurc $\sim$ taxa + depth & \cellcolor{green!15}544862 & \cellcolor{green!40}493808 & \cellcolor{green!45}481869 & \cellcolor{green!50}478851 \\
            \hline
            multifurc $\sim$ taxa * depth & \cellcolor{green!15}544179 & \cellcolor{green!45}489845 & \cellcolor{green!45}481494 & \cellcolor{green!50}478188 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values unknown state information models. \\
          These models are created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values.}
        \label{table:BIC unknown information} 
      \end{table}
      It also follows from this table that the most complex model is the best. In general, the BIC 
        values are smaller than those of the multifurcation models. The modeling here is thus better. \\
      Again, the calculation of finer models (e.g. order or family) was too expensive. \\

      These missing data modeling results can be used to better simulate the data.

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Results of simulation / Influence of different parameters} \label{sec:results - simulation}

    As presented, we compare two methods in our simulation to their prediction accuracy: Fitch and 
      Sankoff. \\
    Figure \ref{fig:influence of unknown data} shows the results of examining various parameters.
    \begin{figure}[h!]
      \centering
      \includegraphics[trim = 0mm 100mm 0mm 0mm, clip, width=\textwidth]{Figures/simulation_evaluation_1.png}
    \end{figure}
    ~ \\ ~ \\ ~ \\
    \begin{figure}
      \centering
      \includegraphics[trim = 0mm 0mm 0mm 150mm, clip, width=\textwidth]{Figures/simulation_evaluation_1.png}
      \caption{Influence of unknown data to prediction. \\
        The first column describes the distributions of free-livings and parasites with a given 
          threshold for the respective simulations to the right. \\
        The middle column investigates the influence of the unknown states, the right the influence of
          the strength of the multifurcation. \\
        The y-axes indicate the percentage of correctly predicted states (including known states). On 
          the x-axis the percentage of forgotten states or missing internal nodes. \\
        Each point corresponds to the average of one hundred simulations, each with 10,000 leaf nodes. \\
        For the middle column we set the strength of the multifurcations to 0.95\% similar to the 
          real data and in the right column the amount of the unknowns to 0.95\% also similar to the 
          real data.}
      \label{fig:influence of unknown data}
    \end{figure}
    First, we tested different transition probability distributions including threshold (first column). 
      It can be observed that the more balanced the percentages of free-livings and parasites, the 
      worse the prediction of the algorithms (second and third column). \\
    It can be seen that both algortithms are always over 50\% prediction and therefore better than 
      guessing. Moreover, they are usually close to each other, with Sankoff always makes better 
      predictions except for equally distributed states as Fitch. \\
    On the other hand, we examined the influence of missing internal nodes (ridge of multifurcation) 
      and missing leaf node information (unknown leaf nodes). \\
    Both factors have a relatively linear influence on the Sankoff method. Fitch, on the other hand, 
      breaks significantly in his prediction from about 70\% unknown leaf nodes or mutifurcation. \\

@@

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Biological view} \label{sec:results - biological view}
      To analyze the results, we have selected some phyla (subtrees) to evaluate our results 
        selectively from the biological point of view: Chordata, Nematoda, Platyhelminthes and 
        Apicomplexa. \\
      In Table \ref{table:phylum leaf nodes states} we compare the given states with the predicted ones. \\
      Several factors such as the distribution of existing input data via parasitics and free-livings, 
        faulty input data from GloBI and reinforcement of errors by multifurcation play an important 
        role, which crystallize out in these examples. \\
      An important factor here is that the credibility of the results. The accuracy of the input data 
        stands and falls with the presence and the correctness of the data of GloBI. Errors of incorrect 
        input data can be amplified by incorrect prediction of unknown species and can be reversed in 
        order to improve the data situation of GloBI. \\
      Since we look at such large trees we can not expect to know all the parasites, so we look at 
        individual positives. This is positive in the sense that the majority have the opposite state.

      \begin{table}
        \begin{center}
          \begin{tabular}{ |l|r||r|r||r|r|r|r| }
            \hline
            & & \multicolumn{2}{c||}{\bfseries original states} & \multicolumn{4}{c|}{\bfseries final states} \\
            \bfseries Phylum & \bfseries \# nodes & \bfseries FL & \bfseries P & \bfseries 0 \bfseries (FL) & \bfseries 0.3 &\bfseries  0.5 & \bfseries 1 (P) \\
            \hline \hline
            Chordata & 91785        & 10451 & 18          & 91759 & 0 & 0 & 26 \\
            &                       & 99.83\% & 0.49\%  & 99.97\% & & & 0.03\% \\ \hline
            Nematoda & 30127        & 21 & 3289           & 1604 & 142 & 1196 & 27185 \\
            &                       & 0.63\% & 99.37\%  & 5.32\% & 0.47\% & 3.97\% & 90.23\% \\ \hline
            Platyhelminthes & 22683 & 7 & 7086            & 175 & & & 22508 \\
            &                       & 0.1\% & 99.9\%    & 0.77\% & 0 & 0 & 99.23\% \\ \hline
            Apicomplexa & 1863      & 1 & 255             & 1 & 0 & 0 & 1862 \\
            &                       & 0.39\% & 99.61\%  & 0.05\% & & & 99.95\% \\
            % \hline \hline
            % Phylum & \# nodes & FL & P
            %   & 0 (FL) & 0.4 & 0.5 & 0.67 & 0.75 & 1 (P) \\
            % Arthropoda & 1198981 & 18912 & 11141 
            %   & 1099509 & 1313 & 22478 & 4176 & 1665 & 70223 \\
            % & & 62.93\% & 37.07\%
            %   & 91.7\% & 0.11\% & 1.87\% & 0.35\% & 0.14\% & 5.86\% \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{Some selected phyla (subtrees), only leaf nodes, are examined for the evolution 
          of the amount and percentage of given data to predicted data.}
        \label{table:phylum leaf nodes states} 
      \end{table}

>>>>

  Here we test in how far it is possible to use the Open Tree of Life (OTL), a synthesis of 
    phylogenetic trees on a backbone taxonomy (resulting in unresolved nodes), to expand available 
    information via phylogenetic trait prediction. We use the Global Biotic Interactions (GloBI) 
    database to categorise 25,962 and 34,860 species as parasites and free-living, respectively and 
    predict states for over $\sim$2.3 million (97.34\%) leaf nodes without state information.

@@

  % This study focuses on the ancestral state reconstruction of parasitism in the tree of life of
  %   Eukaryota. We predict unknown states of species and estimate origins and losses of parasitism. \\
  % The challenge here is the size of the tree and the little information about it. \\
  % Such a large phylogenetic tree does not completely exist and therefore we work with a synthesis 
  %   tree of OTL \cite{Hinchliff2015} which is highly multifurcated. \\
  % For the 2,535,437 leaf nodes we could not gather much data. From the GloBI database 
  %   \cite{Poelen2014} which we used, we could only collect 25,962 parasitic and 34,860 free-living 
  %   species. It follows that we have $\approx 2.4\%$ state information. \\
  % So far, especially small scale or highly manual studies have been carried out. In this scale, it 
  %   requires different data sources to be interconnected. \\
  % We performed an analysis of existing algorithms and selected a Sankoff maximum parsimony algorithm 
  %   using the R package \textit{Castor} \cite{Louca2017}. \\
  % Nevertheless, the results are convincing and even though this is a purely computational approach 
  %   which did not include human experts input, results coincide with prior knowledge. Also regarding 
  %   the number of events, our estimates coincide with previous results by human experts, e.g. the 
  %   study by Weinstein and Kuris \cite{Weinstein2016}. \\

@@

  To accomplish this task, next to the phylogenetic tree information about the current species' 
    states is needed. \\
  Most of the largest interaction databases are offline or outdated. For example: IWDB (Interaction 
    Web Database) \cite{IWDB2003}, Webs on the Web \cite{WOW2004}, Animal Diversity Web 
    \cite{Myers2003} and ecoweb \cite{Cohen2010}. \\
  We use the interaction database Global Biotic Interactions (GloBI) \cite{Poelen2014} because it is 
    including most of the known databases and is still growing actively \cite{Poelen2014}. \\
  The data in GloBI are stored as interactions e.g.  species A parasitize species B. From this we 
    conclude that species A is parasitic and species B free-living. \\
  From this data, we could specify $\sim$2.3 million leaf nodes 34,860 as free-living and 25,962 
    as parasitic.

@@

  The objectives of this work are the following points: (1) Find a suitable ancestral state 
    reconstruction method. (2) Accomplish reconstructing on the Eucaryotic synthesis tree of OTL. \\
  The goal of Point 1 is to evaluate the possible methods based on a simulation of our data 
    situation. \\
  The Sankoff algorithm implemented by Louca et al. is the best in our comparisons. Therefore, 
    point 2 consists of reconstructing the ancestral states and predicting the unknown leaf states 
    with the aid of this algorithm. And then perform an evaluation of the results. \\

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%--------------------------------------------------------------------------------------------------- 
\chapter{Aims}
  The aim of this thesis is the application of maximum parsimony algorithms to non-binary trees and 
    very large data sets to predict unknown leaf node states. \\
  We discuss different ancestral state reconstruction methods and test the most appropriate ones on 
    our data. Hereby, we want to find out how far it is possible to use the Open Tree of Life (OTL) 
    to expandavailable information via phylogenetic trait prediction. As a basis for known states, 
    we use the Global Biotic Interactions (GloBI) database, which cathegorises 34,860 and 25,962 
    species as free-living and parasites. This data is used to perform an ancestral state 
    reconstruction and predict states for over $\sim$2.3 million (97.34\%) leaf nodes without state 
    information.

@@

    Analogous to the missing leaf node states, generalized linear models with binomial regression are 
      compared according to their residuals, based on the "true" / "false" prediction. \\
    Again, in order to compare models of different complexity, the BIC (Bayesian Information Criterion) 
      values are calculated in addition to the residuals. \\
    For all these calculations, the following R functions are used: \textit{glm()}, \textit{anova()} 
      and \textit{BIC()}. \\

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Comparison of different models for multifurcation}
      For the regression analysis of the multifurcation we set up several generalized linear models that 
        could describe multifurcation. \\
      In doing so, we allowed the different influence of the taxa and the heights and depths of a node 
        to be included. From this we got 9 times 4 models of different complexity levels (first row) and 
        the associated BIC values (table \ref{table:BIC multifurcation}).

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & 8273333 & \cellcolor{green!15}7937828 & \cellcolor{green!20}7842157 & \cellcolor{green!30}7644249 \\
            multifurc $\sim$ taxa & 8257680 & \cellcolor{green!15}7922207 & \cellcolor{green!20}7826490 & \cellcolor{green!35}7574154 \\
            \hline
            multifurc $\sim$ taxa + depth & 8273318 & \cellcolor{green!15}7934322 & \cellcolor{green!20}7839364 & \cellcolor{green!35}7539999 \\
            multifurc $\sim$ taxa + max.height & \cellcolor{green!15}7993515 & \cellcolor{green!25}7749121 & \cellcolor{green!30}7661817 & \cellcolor{green!40}7416211 \\
            multifurc $\sim$ taxa + min.height & 8251211 & \cellcolor{green!20}7875521  & \cellcolor{green!25}7778327 & \cellcolor{green!35}7516883 \\
            multifurc $\sim$ taxa + mean.height & \cellcolor{green!20}7825417 & \cellcolor{green!30}7644249 & \cellcolor{green!35}7572474 & \cellcolor{green!45}7340741 \\
            \hline
            multifurc $\sim$ taxa * depth & 8235932 & \cellcolor{green!20}7836755 & \cellcolor{green!25}7757688 & \cellcolor{green!45}7383808 \\
            multifurc $\sim$ taxa * max.height & \cellcolor{green!15}7963438 & \cellcolor{green!30}7693555 & \cellcolor{green!30}7614820 & \cellcolor{green!45}7335338 \\
            multifurc $\sim$ taxa * min.height & 8214030 & \cellcolor{green!20}7808940 & \cellcolor{green!30}7690618 & \cellcolor{green!45}7336627\\
            multifurc $\sim$ taxa * mean.height & \cellcolor{green!25}7768360 & \cellcolor{green!35}7536296 & \cellcolor{green!50}7484953 & \cellcolor{green!50}7206369 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values of the multifurcation models. \\
          These models are created with the R function \textit{glm()} and compared with the \textit{BIC()} 
            function. This results in the listed BIC values, presented in a heatmap colorization.}
        \label{table:BIC multifurcation} 
      \end{table}
      Within every complexity class it can be seen that the mean height gives the best additional factor. \\
      Despite higher complexity, the BIC values are getting smaller from model to model, meaning that 
        the finest model available here is also the best one of these. Lower taxa than orders (e.g. 
        family) are computationally too expensive to calculate. \\
      The model \textit{multifurc $\sim$ order * mean.height} turns out to be the best of our models, 
        whereby it is possible that e.g. \textit{multifurc $\sim$ family * mean.height} is better. \\

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Comparison of different models for missing state information}
      Next to the problem of the multifurcation of the tree is the little interaction data thet we have 
        for the species. For the ancestral state reconstruction, we need information about the states 
        (free-living or parasite) in the leaf nodes. \\
      The eukaryotic synthesis tree has 293,463 leaf nodes. The GloBI database has 5,346,414 interactions 
        (at 29.01.2018). Out of this data we got 51,337 free-living species and 47,332 parasite species 
        for the whole tree of life. From the Eukaryota we could determine 25,962 and 34,860 species as 
        parasites and free-living. With 2,293,463 leaf nodes we still have about 97.34\% unknown leaf 
        nodes.

      We also compared different models in terms of their BICs (Table: \ref{table:BIC unknown information}). \\ 
      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & \cellcolor{green!15}545799 & \cellcolor{green!35}500004 & \cellcolor{green!45}485121 & \cellcolor{green!45}484681 \\
            \hline
            multifurc $\sim$ taxa + depth & \cellcolor{green!15}544862 & \cellcolor{green!40}493808 & \cellcolor{green!45}481869 & \cellcolor{green!50}478851 \\
            \hline
            multifurc $\sim$ taxa * depth & \cellcolor{green!15}544179 & \cellcolor{green!45}489845 & \cellcolor{green!45}481494 & \cellcolor{green!50}478188 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values unknown state information models. \\
          These models are created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values, presented in a heatmap colorization.}
        \label{table:BIC unknown information} 
      \end{table}

      It also follows from this table that the most complex model is the best. In general, the BIC 
        values are smaller than those of the multifurcation models. The modeling here is thus better.
      Again, the calculation of finer models (e.g. family) was too expensive.

      These missing data modeling results could be used to better simulate the data. \\

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Influence of different parameters on the prediction} \label{sec:results - simulation}
    As presented, we compare two methods in our simulation to their prediction accuracy: Fitch and 
      Sankoff.

    First, we tested different distributions of parasites to free-livings including threshold (first 
      column). It can be observed that the more balanced the percentages of free-livings and parasites, 
      the worse the prediction of the algorithms (second and third column). \\
    It can be seen that the predictive power of Sankoff is always greater or equal to the percentage
      of free-living distribution, and therefore more accurate than guessing. \\
    On the other hand, we examined the influence of missing internal nodes (ridge of multifurcation) 
      and missing leaf node information (unknown leaf nodes). \\
    Both factors have a relatively linear influence on the Sankoff method. Fitch, on the other hand, 
      breaks significantly in his prediction from about 70\% unknown leaf nodes or mutifurcation. \\
    Figure \ref{fig:influence of unknown data} shows the results of examining various parameters.

@@

    \begin{figure}[h!]
      \centering
      \includegraphics[trim = 0mm 100mm 0mm 0mm, clip, width=\textwidth]{Figures/simulation_evaluation_1.png}
    \end{figure}
    ~ \\
    \begin{figure}[h!]
      \centering
      \includegraphics[trim = 0mm 0mm 0mm 150mm, clip, width=\textwidth]{Figures/simulation_evaluation_1.png}
      \caption{Influence of unknown data to prediction. \\
        The first column describes the distributions of free-livings and parasites with a given 
          threshold for the respective simulations to the right. Note, these distributions are chosen 
          to form the specified proportions with the threshold, ignoring the number of state changes. \\
        The middle column investigates the influence of the unknown states, the right the influence of
          the strength of the multifurcation. \\
        The y-axes indicate the percentage of correctly predicted states (including known states). On 
          the x-axis the percentage of forgotten states or missing internal nodes. \\
        Each point corresponds to the average of one hundred simulations, each with 10,000 leaf nodes. \\
        For the middle column we set the strength of the multifurcations to 0.95\%, similar to the 
          real data and in the right column the amount of the unknowns to 0.95\% also similar to the 
          real data.}
      \label{fig:influence of unknown data}
    \end{figure}

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Biological view} \label{sec:results - biological view}
      To analyze the results, we have selected some phyla (subtrees) to evaluate our results 
        selectively from the biological point of view: Chordata, Nematoda, Platyhelminthes and 
        Apicomplexa. \\
      In Table \ref{table:phylum leaf nodes states} we compare the given states with the predicted ones. \\
      Several factors such as the distribution of existing input data via parasitics and free-livings, 
        faulty input data from GloBI and reinforcement of errors by multifurcation play an important 
        role, which crystallize out in these examples. \\
      An important factor here is the credibility of the results. The accuracy of the input data 
        stands and falls with the presence and the correctness of the data of GloBI. Errors of incorrect 
        input data can be amplified by incorrect prediction of unknown species and can be reversed in 
        order to improve the data situation of GloBI. \\
      Since we look at such large trees we can not expect to know all the parasites, so we look at 
        individual positives. This is positive in the sense that the majority have the opposite state.

      \newcommand*\rot{\rotatebox{90}}

      % \tiny	\scriptsize \footnotesize \small \normalsize \large \Large \LARGE \huge \Huge
      \begin{table}
        \begin{center}
          \hspace*{-0.5cm}\begin{tabular}{|>{\scriptsize}c|>{\scriptsize}l|>{\scriptsize}r|>{\scriptsize}r||>{\scriptsize}r|>{\scriptsize}r||>{\scriptsize}r|>{\scriptsize}r|>{\scriptsize}r||>{\scriptsize}r|>{\scriptsize}r|>{\scriptsize}r| }
            \hline
            & \bfseries Taxa & \bfseries \#internal & \bfseries \#leaf 
                & \multicolumn{2}{>{\scriptsize}c||}{\bfseries known states} & \multicolumn{3}{>{\scriptsize}c|}{\bfseries final states}
                & \bfseries Root & \bfseries \#origins & \bfseries \#losses \\
            & \bfseries subtree & \multicolumn{2}{>{\scriptsize}c||}{\bfseries nodes} & \bfseries FL & \bfseries P
                & \bfseries [0, 0.5) & \bfseries 0.5 & \bfseries (0.5, 1]
                & \bfseries node & \multicolumn{2}{>{\scriptsize}c|}{\bfseries (with rounding)} \\
            & \bfseries & & & & 
                & \bfseries FL & & \bfseries P 
                & state & \bfseries FL $\rightarrow$ P & \bfseries P $\rightarrow$ FL \\
            \hline \hline
            \parbox[c][7mm][c]{1mm}{\multirow{2}{*}{\rot{\bfseries Domain~}}}
            & Eukaryota & 241974 & 2293463      & 34860 & 25962     & 2053212 & 7775 & 474450     & 1 & 462 & 369 \\
            & & &                               & 57.31\% & 42.69\% & 80.98\% & 0.31\% & 18.71\%  & P & & \\
            \hline \hline
            \parbox[c]{1mm}{\multirow{6}{*}{\rot{\bfseries Kingdom}}}
            & Chloro- & 43486 & 416478          & 3519 & 77         & 410795 & 4182 & 1501        & 0.5 & 97 & 222 \\
            & plastida & &                      & 97.86\% & 2.14\%  & 98.63\% & 1.00\% & 0.36\%   & & & \\ \cline{2-12}
            & Fungi & 9534 & 31457              & 577 & 2983        & 38520 & 5723 & 266463       & 0 & 42 & 2 \\
            & & &                               & 16.21\% & 83.79\% & 12.40\% & 1.84\% & 85.76\%  & FL & & \\ \cline{2-12}
            & Metazoa & 179944 & 1491012        & 30758 & 22373     & 1329065 & 25535 & 136412    & 0.5 & 321 & 129 \\
            & & &                               & 57.89\% & 42.11\% & 89.14\% & 1.71\% & 9.15\%   & & & \\
            \hline \hline
            \parbox[c]{1mm}{\multirow{10}{*}{\rot{\bfseries Phylum}}}
            & Apicomplexa & 239 & 1863          & 1 & 255           & 1 & 0 & 1862                & 1 & 0 & 1 \\
            & & &                               & 0.39\% & 99.61\%  & 0.05\% & 0\% & 99.95\%      & P & & \\ \cline{2-12}
            & Arthropoda & 120479 & 1198981     & 18912 & 11141     & 1100822 & 22478 & 76064     & 0 & 281 & 108 \\
            & & &                               & 62.93\% & 37.07\% & 91.78\% & 1.87\% & 6.34\%   & FL & & \\ \cline{2-12}
            & Chordata & 30761 & 91785          & 10451 & 18        & 91759 & 0 & 26              & 0 & 12 & 1 \\
            & & &                               & 99.83\% & 0.49\%  & 99.97\% & 0\% & 0.03\%      & FL & & \\ \cline{2-12}
            & Nematoda & 3437 & 30127           & 21 & 3289         & 1746 & 1196 & 27185         & 1 & 2 & 11 \\
            & & &                               & 0.63\% & 99.37\%  & 5.79\% & 3.97\% & 90.23\%   & P & & \\ \cline{2-12}
            & Platyhel- & 4459 & 22683       & 7 & 7086          & 175 & 0 & 22508             & 0 & 0 & 5 \\
            & minthes & &                          & 0.1\% & 99.9\%    & 0.77\% & 0\% & 99.23\%      & P & & \\
            \hline \hline 
            \parbox[c]{1mm}{\multirow{2}{*}{\rot{\bfseries Class~}}}
            & Insecta & 91256 & 989572          & 17841 & 10734     & 1022747 & 976 & 57105       & 0 & 245 & 77 \\
            & & &                               & 62.44\% & 37.56\% & 94.63\% & 0.09\% & 5.28\%   & FL & & \\ 
            \hline
          \end{tabular} 
        \end{center}
        \caption{Some selected phyla (subtrees), only leaf nodes, are examined for the evolution 
          of the amount and percentage of given data to predicted data.}
          % \caption{All kingdoms (subtrees), only leaf nodes, are examined for the evolution of the 
          % amount and percentage of given data to predicted data.}
          % \caption{Origins and losses}
        \label{table:phylum leaf nodes states} 
      \end{table}
